Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-02-13 Paper.Rmd
excised TL from main text

We tested whether the linearity of the spatial Taylor's law relationship was partly
due to the copula structure of the dependences between the populations sampled. 
Using corn leaf aphid abundance data as an example, but also
applying the same analysis to the abundance data we had for 19 other aphid species
(see *Data*), we calculated the curvature (details in Appendix \ref{SM-TL_curvature}) 
of the empirical Taylor's law relationship; Taylor's law hypothesizes that this 
curvature should be close to 0. We then calculated the same curvature quantity for
10000 surrogate datasets, using the same surrogate procedure described above in 
this section, and compared the empirical Taylor's law curvature to the 
distribution of surrogate curvatures using a two-tailed test. If the empirical
curvature value was below the $5^{\text{th}}$ quantile of the surrogate distribution, 
then copula structure contributes to the linearity of spatial Taylor's law 
because randomizing copula structure while retaining other statistical properties
of the data cause the Taylor's law relationship to become significantly more 
curved.

<!--***DAN: Shya,  want to add the extinction risk stuff back in, let's discuss-->

@@

```{r read_tl_aphid_count_sp6, echo=F}
tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
tl_aphid_count<-readRDS("./Results/taylorslaw_results/tl_aphid_count.RDS")
aphid_info<-readRDS("./Results/aphid_info.RDS")
pleft<-tl_aphid_count$p_left_mncurv
stat<--2*sum(log(pleft))
pchi2<-1-pchisq(stat,40)
idsp<-which(pleft<0.05)
pleftgood<-pleft[idsp]
```

Results also suggested that copula structure contributes to the linearity of 
spatial Taylor's law. For instance, the spatial Taylor's law relationship 
for corn leaf aphid abundance data was significantly less curved than for 
surrogate datasets (Fig. \ref{fig_sTL_mncurv_sp6}). Three of $20$ aphid species
had cuvatures less than at least $95\%$ of surrogate datasets ($p=$ `r pleftgood` for 
the `r aphid_info$c1[idsp]` species, respectively). 
<!--***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)
***Shya: Done.-->
Three such significant
results of 20 species could have occurred by chance due to 
multiple-testing error. However, for each of the 20 species, we computed the fraction
of surrogate curvatures less than the curvature that came from real data, and we
combined these $p$-values using Fisher's combined probability test to get an
aggregate $p$-value of `r pchi2`
<!--***DAN: Shya, see the Wikipedia article on Fisher's method to learn about it. Then 
you need to put this analysis into the code suite and have the resulting p value
typed in above be an autolink instead. Here is what I did: 
1) I copied by hand all the pleft values on the pdf you sent me into a vector pleft. Instead you should create such a vector in your code from your test output, not typing in the results.
2) I used the formula stat<--2*sum(log(pleft)) to get the Fisher method statistic
3) I compared to a chi-squared distribution with 40 degress of freedom using 1-pchisq(stat,40) to get my p-value
Please let me know if any of this is unclear.
***Shya: Done.-->
for the test of the null hypothesis that 
copula structure did not contribute to linearity of Taylor's law for any of the 
species. This result is usefully suggestive and can guide future work, but still does not
provide certainty, because the Fisher method provides $p$-values that are too 
small if its assumption of independent tests is violated. Though our 20 species
have different host plants so may be independent for the purposes of our analyses, 
it would be difficult to fully assess independence.
<!--***DAN: Shya, can you remind me why the same analyses were not done for plankton,
if there is a good reason? Or if you did it, what were the results?
***Shya : we did it, I have the pdf but I don't have the code where I called to generate the pdf,
As we could not get any significant results for a single species (see attachment in the mail I sent you)
we did not show the results.--> 
We provide simulation results in Appendix \ref{SM-Consequence_TL_simulation}
that robustly illustrate, at least in the model setup we used, 
that copula structure influences spatial Taylor's law.

\begin{figure}[!h]
\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
\caption{Mean curvature of the spatial Taylor's law relationship for 
corn leaf aphid abundance
data (A) was significantly less than for surrogates (B, C). The Taylor's law 
relationship for one example surrogate (B) shows shows greater curvature than
for data (A), and the significance of the difference was demonstrated
by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
\end{center}
\end{figure}
<!--***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
***sHYA: dONE.-->

<!--***DAN: I want to put back in some extinction risk results. Let's talk about this, please, Shyamolina-->

>>>>

<!--***DAN: 
1) Shya put concepts on why we would expect lt dependence of noise to mean
more ext risk in a metapop context
2) then methods for how we test that (present the model, say what sims you did-->

@@

<!--***DAN: Shya to insert extinction risk results with one fig-->

### 2019-02-13 SupportingInformation.Rmd
excised TL from main text

>>>>

<!--
Stuff on TL cut from the main text
From Q3 methods:
We tested whether the linearity of the spatial Taylor's law relationship was partly
due to the copula structure of the dependences between the populations sampled. 
Using corn leaf aphid abundance data as an example, but also
applying the same analysis to the abundance data we had for 19 other aphid species
(see *Data*), we calculated the curvature (details in Appendix \ref{SM-TL_curvature}) 
of the empirical Taylor's law relationship; Taylor's law hypothesizes that this 
curvature should be close to 0. We then calculated the same curvature quantity for
10000 surrogate datasets, using the same surrogate procedure described above in 
this section, and compared the empirical Taylor's law curvature to the 
distribution of surrogate curvatures using a two-tailed test. If the empirical
curvature value was below the $5^{\text{th}}$ quantile of the surrogate distribution, 
then copula structure contributes to the linearity of spatial Taylor's law 
because randomizing copula structure while retaining other statistical properties
of the data cause the Taylor's law relationship to become significantly more 
curved.

From Q3 results:
```{r read_tl_aphid_count_sp6, echo=F}
tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
tl_aphid_count<-readRDS("./Results/taylorslaw_results/tl_aphid_count.RDS")
aphid_info<-readRDS("./Results/aphid_info.RDS")
pleft<-tl_aphid_count$p_left_mncurv
stat<--2*sum(log(pleft))
pchi2<-1-pchisq(stat,40)
idsp<-which(pleft<0.05)
pleftgood<-pleft[idsp]
```

Results also suggested that copula structure contributes to the linearity of 
spatial Taylor's law. For instance, the spatial Taylor's law relationship 
for corn leaf aphid abundance data was significantly less curved than for 
surrogate datasets (Fig. \ref{fig_sTL_mncurv_sp6}). Three of $20$ aphid species
had cuvatures less than at least $95\%$ of surrogate datasets ($p=$ `r pleftgood` for 
the `r aphid_info$c1[idsp]` species, respectively). 
***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)
***Shya: Done.
Three such significant
results of 20 species could have occurred by chance due to 
multiple-testing error. However, for each of the 20 species, we computed the fraction
of surrogate curvatures less than the curvature that came from real data, and we
combined these $p$-values using Fisher's combined probability test to get an
aggregate $p$-value of `r pchi2`
***DAN: Shya, see the Wikipedia article on Fisher's method to learn about it. Then 
you need to put this analysis into the code suite and have the resulting p value
typed in above be an autolink instead. Here is what I did: 
1) I copied by hand all the pleft values on the pdf you sent me into a vector pleft. Instead you should create such a vector in your code from your test output, not typing in the results.
2) I used the formula stat<--2*sum(log(pleft)) to get the Fisher method statistic
3) I compared to a chi-squared distribution with 40 degress of freedom using 1-pchisq(stat,40) to get my p-value
Please let me know if any of this is unclear.
***Shya: Done.
for the test of the null hypothesis that 
copula structure did not contribute to linearity of Taylor's law for any of the 
species. This result is usefully suggestive and can guide future work, but still does not
provide certainty, because the Fisher method provides $p$-values that are too 
small if its assumption of independent tests is violated. Though our 20 species
have different host plants so may be independent for the purposes of our analyses, 
it would be difficult to fully assess independence.
***DAN: Shya, can you remind me why the same analyses were not done for plankton,
if there is a good reason? Or if you did it, what were the results?
***Shya : we did it, I have the pdf but I don't have the code where I called to generate the pdf,
As we could not get any significant results for a single species (see attachment in the mail I sent you)
we did not show the results. 
We provide simulation results in Appendix \ref{SM-Consequence_TL_simulation}
that robustly illustrate, at least in the model setup we used, 
that copula structure influences spatial Taylor's law.

\begin{figure}[!h]
\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
\caption{Mean curvature of the spatial Taylor's law relationship for 
corn leaf aphid abundance
data (A) was significantly less than for surrogates (B, C). The Taylor's law 
relationship for one example surrogate (B) shows shows greater curvature than
for data (A), and the significance of the difference was demonstrated
by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
\end{center}
\end{figure}
***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
***sHYA: dONE.
-->
